\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-08 casestudies.tex
conclusion
shortening
edits

To demonstrate the applicability of our design and implementation, we
conduct three case studies.

@@

Elimination of inductive types defined via \lsti{FInductive} is
mostly via the \lsti{FRecursion} and \lsti{FInduction} commands.
An exception is a handful of trivial ``inversion lemmas''.
For example, consider the lemma \lsti!\forall t, \neg step tm_true t!
stating that \lsti!tm_true! is irreducible.
If \lsti{step} were an ordinary inductive type, then it could be proved in Coq
simply by \lsti{intros t H; inversion H}.
But since \lsti{step} is extensible, it seems that the lemma has to be
proved by \lsti{FInduction} on \lsti{step}, and that the programmer
has to verify that a derived family does not accidentally make \lsti{tm_true} reducible.
%lemma still holds in derived families that add new constructors to \lsti{step}.
%
We observe that it is lighter-weight to use overriding (\cref{sec:override}) instead:
the programmer can specify that the proof of the lemma should be overridden
in any derived family that further binds \lsti{step},
and in return, they are permitted to treat \lsti{step} as an ordinary
inductive type in the proof and thus discharge the proof with \lsti{inversion}.
The plugin then automatically tries the same proof script in a derived
family to override the proof. Although proof scripts rather
than proof terms are reused, rechecking of proof terms seems justified
by the triviality of the lemmas and the terseness of the proofs.

We note that the current plugin implementation does not support
mutually inductive types or mutual induction, which will be
useful for modeling languages much more complex than STLCs.
We believe that generalizing our design to mutually inductive types does
not pose theoretical challenges, however.

>>>>

%To demonstrate the applicability of our design and implementation, we
%conduct two case studies.

@@

Elimination of inductive types defined via \lsti{FInductive} is
mostly via the \lsti{FRecursion} and \lsti{FInduction} commands.
An exception is a handful of trivial ``inversion lemmas''.
For example, consider the lemma \lsti!\forall t, \neg step tm_true t!
stating that \lsti!tm_true! is irreducible.
If \lsti{step} were an ordinary inductive type, then it could be proved in Coq
simply by \lsti{intros t H; inversion H}.
But \lsti{step} is extensible. So one way to prove the lemma is
by \lsti{FInduction} on \lsti{step} and verifying that
a derived family does not accidentally make \lsti{tm_true} reducible.
%lemma still holds in derived families that add new constructors to \lsti{step}.
%
We observe that it is lighter-weight to use overriding (\cref{sec:override}) instead:
the programmer can specify that the proof of the lemma should be overridden
in any derived family that further binds \lsti{step},
and in return, they are permitted to treat \lsti{step} as an ordinary
inductive type in the proof and thus use \lsti{inversion} to prove it.
The plugin then automatically tries the same proof script in a derived
family to override the proof. Although proof scripts rather
than proof terms are reused, the practice seems justified
by the triviality of the lemmas and the terseness of the proof scripts.

The current plugin implementation does not yet support
mutually inductive types or mutual induction, which are
useful for modeling languages much more complex than STLCs.
However,
we believe generalizing our design to mutually inductive types does
not pose theoretical challenges.

### 2022-11-08 metatheory-condensed.tex
edits

\TT is intended as a foundational model. So unlike our programmer-facing plugin,
typing is structural rather than nominal, and
partial recursors and propositional equalities are not automated but have to
be explicitly defined.
\TT provides MLTT-style constructs that can be used to \emph{encode}
family polymorphism rather than building it in directly.
%
Most notably, \TT extends MLTT with what we call \emph{linkages}.\footnote{
Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
family polymorphism in an OO setting, but the technical details differ significantly.}
Linkages model families, so they are like tuples but with a twist: they
support late binding.
Unlike dependent tuples where a later component is \emph{existentially}
quantified over the earlier ones, a linkage component is \emph{universally}
quantified over---and thus polymorphic to---components preceding it.
%
\TT also features \emph{linkage transformers} that can be used to inductively
construct linkages from other linkages, by inheriting or overriding components of a linkage,
or by extending linkages with new components.
%
Inductive types are modeled as W-types~\cite{martin1984intuitionistic} and their
extension as overriding.
%
Consistency and canonicity are proved by giving semantic models to the syntax.

>>>>

\TT is intended as a foundational model. So unlike our programmer-facing plugin,
typing is structural rather than nominal, and
partial recursors and propositional equalities are not automated but have to
be explicitly defined.
\TT provides MLTT-style constructs that can be used to \emph{encode}
family polymorphism rather than building it in directly.
%
Most notably, \TT extends MLTT with what we call \emph{linkages}.\footnote{
Linkages are a namesake of the theoretical device through which prior work~\cite{zm2017} models
family polymorphism in an OO setting, but the technical details differ significantly.}
Linkages model families, so they are like tuples but with a twist: they
support late binding.
Unlike dependent tuples where a later component is \emph{existentially}
quantified over the earlier ones, a linkage component is \emph{universally}
quantified over---and thus polymorphic to---components preceding it.
%
\TT also features \emph{linkage transformers} that can be used to inductively
construct linkages from other linkages, by inheriting or overriding components of a linkage,
or by extending a linkage with new components.
%
Inductive types are modeled as W-types~\cite{martin1984intuitionistic} and their
extension as overriding.
%
Consistency and canonicity are proved by giving semantic models to the syntax.
